---
title:  Assignment  3
author: |
  | Last name: Du
  | First name: Min
  | Student ID: 1002602230
output: pdf_document
header-includes: \usepackage{color,amsmath}
fontsize: 12pt
---
#\textcolor{blue}{Part 1}
1.
Applied Chemistry.

2.
Schnurr, P. J., Molenda, O., Edwards, E., Espie, G. S., & Allen, D. G. (2016). Improved biomass productivity in algal biofilms through synergistic interactions between photon flux density and carbon dioxide concentration. Bioresource Technology, 219, 72-79.

3.
Department of Chemical Engineering.

4.
https://doi.org/10.1016/j.biortech.2016.06.129

5.
R (1.15.1), StatPlus.

6.
From experiment.

7.
Yes, the author presents a summary statistics of statistically significant coefficients that describe surface response model used to determine the impact of [CO2]*[PFD] on biomass productivity.

8.
Yes, the authors use tests. For example, Biofilm biomass productivities at various CO2 concentrations were determined by linear regression analysis (95% confidence intervals) of plots of biofilm yields across the time course of the experiment.

9.
The values were reported to 2 decimal places.

10.
Linear regression analysis.



\newpage
#\textcolor{blue}{Part 2}
\textbf{Question 1}

(a)

```{r echo=FALSE, message=F, warning=F, cache=T}
video2230=read.csv("/Users/mindu/Desktop/STA303/Assignment3/video.csv")
attach(video2230)
sex_like2230=table(video2230$sex,video2230$like)
sex_like2230
```

```{r echo=FALSE, message=F, warning=F, cache=T}
chisq.test(sex_like2230,correct = FALSE)
```

$H_0$ = Sex and Student's preference for playing video games are independent.

$H_a$ = Sex and Student's preference for playing video games are not independent.

First P-value = 0.0680, which is smaller than significance level 0.1, reject $H_0$, there is evidence of association between sex and student's preference for playing video games.

```{r echo=FALSE, message=F, warning=F, cache=T}
fisher.test(sex_like2230)
```

$H_0$ = Sex and Student's preference for playing video games are independent.

$H_a$ = Sex and Student's preference for playing video games are not independent.

Second P-value = 0.0782, smaller than significance level 0.1, reject $H_0$, there is evidence of association between sex and student's preference for playing video games.

(b)


```{r echo=FALSE, message=F, warning=F, cache=T}
gradeA_2230= table(video2230$sex[video2230$grade == "A"], video2230$like[video2230$grade == "A"])
gradeA_2230
chisq.test(gradeA_2230,correct = F)
```

$H_0$ = Sex and a student's preference for grade A are independent.

$H_a$ = Sex and a student's preference for grade A are not independent.

When grade is A, p-value is 0.0011, which is smaller than 0.1,reject $H_0$, Sex and a student's preference for grade A are not independent at 0.1 significance level.

```{r echo=FALSE, message=F, warning=F, cache=T}
gradenA_2230 = table(video2230$sex[video2230$grade == "nA"], video2230$like[video2230$grade == "nA"])
gradenA_2230
chisq.test(gradenA_2230, correct = F)
```

$H_0$ = Sex and a student's preference for grade nA are independent.

$H_a$ = Sex and a student's preference for grade nA are not independent.

When grade is nA, p-value is 0.9421, which is greater than 0.1, do not reject $H_0$, Sex and a student's preference for grade nA are independent at 0.1 significance level.

Therefore, the association between sex and students' preference for playing video games change with grade expected.



\textbf{Question 2}
(a) 

model2.1: $log(\frac{\pi}{1-\pi})$=$\beta_0+\beta_1I_{male}+\beta_2I_{gradenA}+\beta_3I_{male}*I_{gradenA}$

model2.2: $log(\frac{\pi}{1-\pi})$=$\beta_0+\beta_1I_{male}+\beta_2I_{gradenA}$

$I_{male} \{^{=1, Male}_{=0,Female}$

$I_{gradenA}\{^{=1, gradenA}_{=0,gradeA}$

$I_{male}*I_{gradenA}\{^{=1, male,gradenA}_{=0,Otherwise}$

Test1:

$H_0$: the coefficient of interaction term is 0. ($\beta_3=0$)

$H_a$: the coefficient of interaction term is not 0. ($\beta_3\ne0$)

```{r echo=FALSE, message=F, warning=F, cache=T}
model21_2230 =glm(like~sex+grade+sex * grade, family = binomial(link = "logit") ,data = video2230)
model22_2230 =glm(like~sex+grade, family = binomial(link = "logit") ,data = video2230)
anova(model21_2230,model22_2230,test="Chisq")
```

P-value=0.0087, smaller than 0.1, reject $H_0$, there is significant interaction effect, model 2.1 is better.

Test2:

$H_0$: the coefficient of interaction term is 0. ($\beta_3=0$)

$H_a$: the coefficient of interaction term is not 0. ($\beta_3\ne0$)

```{r echo=FALSE, message=F, warning=F, cache=T}
summary(model21_2230)
```

In summary, P-value of interaction term is 0.0185 < 0.1, reject $H_0$, the interaction term is significant, model2.1 includes interaction term is better.

Two tests give the same result that there exists interaction effect. Therefore, model2.1 includes interaction term is better.

(b)

There is significant interaction effect between sex and grade. The probability of students' preference for playing video games between male and female differ with grade type. 

In this context, Sex effect is affected by grade, which is consistent with part 1(b), so they are agree.

\textbf{Question 3}
(a)
```{r echo=FALSE, message=F, warning=F, cache=T}
count2230=table(video2230)
count_2230=as.data.frame(count2230)
colnames(count_2230)[4]= "count"
count_2230
``` 


model3.1: log($\mu_{ijk}$) ~ $\beta_0+\beta_{1}I_{[sex=male]} + \beta_{2}I_{[grade=nA]} + \beta_{3}I_{[like=yes]} + \beta_{4}I_{[sex=male]}*I_{[grade=nA]} + \beta_{5}I_{[sex=male]}*I_{[like=yes]}  + \beta_{6}I_{[grade=nA]}*I_{[like=yes]} + \beta_{7}I_{[sex=male]}*I_{[grade=nA]}*I_{[like=yes]}$

model3.2: log($\mu_{ijk}$) ~ $\beta_0+\beta_{1}I_{[sex=male]} + \beta_{2}I_{[grade=nA]} + \beta_{3}I_{[like=yes]} + \beta_{4}I_{[sex=male]}*I_{[grade=nA]} + \beta_{5}I_{[sex=male]}*I_{[like=yes]}  + \beta_{6}I_{[grade=nA]}*I_{[like=yes]}$

$I_{[sex=male]} \{^{=1, Male}_{=0,Female}$

$I_{[grade=gradenA]}\{^{=1, gradenA}_{=0,gradeA}$

$I_{[like=yes]}\{^{=1, like=yes}_{=0,like=no}$

$I_{male}*I_{gradenA}\{^{=1, male,gradenA}_{=0,Otherwise}$

$I_{male}*I_{yes}\{^{=1, male,yes}_{=0,Otherwise}$

$I_{gradenA}*I_{yes}\{^{=1, gradenA,yes}_{=0,Otherwise}$

$I_{[sex=male]}*I_{[grade=nA]}*I_{[like=yes]}\{^{=1, male,gradenA,yes}_{=0,Otherwise}$

(b)

i.

```{r message=F, warning=F, cache=T, include=FALSE}
model31_2230 = glm(count ~  sex + grade + like + sex*grade + sex*like + grade*like + sex*grade*like, family = poisson() , data = count_2230)
model32_2230 = glm(count ~  sex + grade + like + sex*grade + sex*like + grade*like, family =  poisson() , data = count_2230)
deviance(model31_2230)
deviance(model32_2230)
deviance(model21_2230)
deviance(model22_2230)
```

Model    | Deviance
:--------|:--------------:|
model31  | -5.773162e-15
model32  | 6.878764
model21  | 85.15215
model22  | 92.03091

Deviance for poisson models are much lower.

ii.

```{r message=F, warning=F, cache=T, include=FALSE}
library(aod)
wald.test(Sigma = vcov(model31_2230),b=coef(model31_2230),Term = 8)
```

Model                | Test Statistics          |Distribution             |
:--------------------|:------------------------:|:-----------------------:|
Logistic Regression  | 6.8788-(-5.7732e-15)=6.87|$X^2(1)$                 |
Poisson Regression   | 92.031-85.152=6.87       |$X^2(1)$                 |


Note that the distribution for using wald test in part2 and and part3 is the same, follows $X^2(1)$.

The Test Statistics in part2 under logistic regression is 6.8788-(-5.7732e-15)=6.87.

The Test Statistics in part3 under poisson distribution is 92.031-85.152=6.87.

Two test statistics are the same. 

According to Wald test,$H_0:\beta_{7}=0$, $H_a:\beta_{7}\ne0$, p-value is 0.018, smaller than significance level 0.1, reject $H_0$, $\beta_{7}\ne0$.

Under the same distribition and the same test statistics. The wald test gives the same result.


iii.

In poisson regression, increase $x_{ijk}$ by one unit, holding other predictors constant.$\mu_j$ changes by a factor of $e^{\beta_{ijk}}$, poisson regression focus on investigating the count affected by predictors.

In logistic model, increase $x_{ijk}$ by one unit, holding other predictors constant.$\mu_j$ changes log odd ratio by $\beta$, logistic model focus on investigating the probability of like affected by predictors.

\newpage
# \textcolor{blue}{Appendix}
# \textbf{(Part 2)}
1.
(a)
```{r,eval=FALSE}
video2230=read.csv("/Users/mindu/Desktop/STA303/Assignment3/video.csv")
attach(video2230)
sex_like2230=table(video2230$sex,video2230$like)
sex_like2230
```

```{r,eval=FALSE}
chisq.test(sex_like2230,correct = FALSE)
```

```{r,eval=FALSE}
fisher.test(sex_like2230)
```

(b)
```{r,eval=FALSE}
gradeA_2230= table(video2230$sex[video2230$grade == "A"], video2230$like[video2230$grade == "A"])
gradeA_2230
chisq.test(gradeA_2230,correct = F)
gradenA_2230 = table(video2230$sex[video2230$grade == "nA"], video2230$like[video2230$grade == "nA"])
gradenA_2230
chisq.test(gradenA_2230, correct = F)
```

2.
(a) 
Test1:
```{r,eval=FALSE}
model21_2230 =glm(like~sex+grade+sex * grade, family = binomial(link = "logit") ,data = video2230)
model22_2230 =glm(like~sex+grade, family = binomial(link = "logit") ,data = video2230)
anova(model21_2230,model22_2230,test="Chisq")
```

Test2:
```{r,eval=FALSE}
summary(model21_2230)
```

3.
(a)
```{r,eval=FALSE}
count2230=table(video2230)
count_2230=as.data.frame(count2230)
colnames(count_2230)[4]= "count"
count_2230
```

(b)
i.
```{r,eval=FALSE}
model31_2230 = glm(count ~  sex + grade + like + sex*grade + sex*like + grade*like + sex*grade*like, family = poisson() , data = count_2230)
model32_2230 = glm(count ~  sex + grade + like + sex*grade + sex*like + grade*like, family =  poisson() , data = count_2230)
deviance(model31_2230)
deviance(model32_2230)
deviance(model21_2230)
deviance(model22_2230)
```

ii.
```{r,eval=FALSE}
summary(model31_2230)
summary(model32_2230)
summary(model21_2230)
summary(model22_2230)
library(aod)
wald.test(Sigma = vcov(model31_2230),b=coef(model31_2230),Term = 8)
```
